---
title: "FA and Variances"
author: "David Gerard"
date: "September 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I show that the variance estimates are the same whether we use Wang et al's approach or Gagnon-Bartsch et al's approach.

# Simulate Data from Assumed Model

```{r}
set.seed(1)
n <- 11
p <- 103
q <- 3
k <- 2
pi0 <- 0.8
X <- model.matrix(~sample(x = c(0, 1), size = n, replace = TRUE, prob = c(1/2, 1/2)))
colnames(X) <- c("Intercept", "Treatment")
Beta <- matrix(rnorm(k * p), nrow = k)
which_null <- sample(x = c(TRUE, FALSE), size = p, replace = TRUE, prob = c(pi0, 1 - pi0))
Beta[2, which_null] <- 0
Z <- X %*% matrix(rnorm(k * q), nrow = k) + matrix(rnorm(n * q), nrow = n)
Alpha <- matrix(rnorm(q * p), nrow = q)
sig2_diag_true <- rchisq(n = p, df = 4) / 4
E <- matrix(rnorm(n * p), nrow = n) %*% diag(sqrt(sig2_diag_true))
Y <- X %*% Beta + Z %*% Alpha + E
```



# Get the two different matrices
```{r}
Q <- qr.Q(qr(X), complete = TRUE)
Q2 <- Q[, -c(1:2), drop = FALSE]
Y3 <- crossprod(Q2, Y) # The Wang matrix
Resid_mat <- Q2 %*% Y3 # The Gagnon-Bartsch Matrix
```

# Sanity check that `Resid_mat` is actually the residuals
```{r}
Resid2 <- Y - X %*% solve(t(X) %*% X) %*% t(X) %*% Y
plot(c(Resid_mat), c(Resid2))
abline(0, 1, col = 2, lty = 2)
```

# Do FA on both

```{r}
library(vicar)
wout <- pca_naive(Y = Y3, r = q)
gout <- pca_naive(Y = Resid_mat, r = q)

plot(abs(c(wout$alpha)), abs(c(gout$alpha)), main = "alpha for wang vs for gb")
abline(0, 1, lty = 2)
```

The alpha estimates are slightly different because we scale them by `sqrt(nrow(Y))`, where `nrow(Y)` is `r n` in the residual matrix and `r n - k` in the rotated matrix.

```{r}
plot(wout$sig_diag, gout$sig_diag, main = "wang vs gb variances")
abline(0, 1)
```

The same thing here, we divide the residual sums of squares by `nrow(Y) - r`, which is `r nrow(Y) - q` in the Gagnon-Bartsch matrix and `r nrow(Y) - q - k` in the Wang matrix.

However, I would imagine that any sane person would divide by `nrow(Y) - k - q` regardless because we already regressed out the observed covariates. If so, then the variance estimates are the exact same.

```{r}
plot(wout$sig_diag, gout$sig_diag * (n - q) / (n - q - k), main = "wang vs (adjusted) gb variances")
abline(0, 1)
```

# ML method
```{r}
w_ml <- fa_ml(Y = Y3, r = q)
g_ml <- fa_ml(Y = Resid_mat, r = q)
```

For the MLE FA, it doesn't seem to work out as exactly.

```{r}
plot(abs(c(w_ml$alpha)), abs(c(g_ml$alpha)))
abline(0, 1, lty = 2, col = 2)
```

```{r}
plot(w_ml$sig_diag, g_ml$sig_diag, main = "wang vs gb variances")
abline(0, 1)

plot(w_ml$sig_diag, g_ml$sig_diag * (n - q) / (n - k - q), main = "wang vs gb (adjusted) variances")
abline(0, 1)
```


```{r}
sessionInfo()
```
