---
title: "Compare MOUTHWASH Factor with Chris' Factor"
author: "David Gerard"
date: "January 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Using the notation from Gerard and Stephens (2017), the second step of MOUTHWASH uses the model
$$
\hat{\beta} \sim N_p(\beta + \hat{\alpha}^Tz, \xi \Sigma / \|x\|^2),
$$
where $\Sigma = diag(\hat{\sigma}_1^2,\ldots,\hat{\sigma}_p^2)$. In Chris' write-up, he proposes fitting
$$
\hat{\beta} \sim N_p(\beta + \hat{\alpha}^Tz, [1/\|x\|^2 + \|z\|^2 / (n-1)]\Sigma).
$$
As a first step to see if Chris' idea is correct, we will see if the variance inflation factors approximately equal each other (or are at least correlated):
$$
\xi/\|x\|^2 \approx 1/\|x\|^2 + \|z\|^2 / (n-1)\\
\Leftrightarrow \xi \approx 1 + \|x\|^2\|z\|^2 / (n-1).
$$

This is easy to check using the current implementation of MOUTHWASH.

# Analysis

```{r, message=FALSE, warning=FALSE, eval=FALSE}
library(vicar)
library(seqgendiff)
library(sva)
hdat <- as.matrix(read.csv("../data/heart.csv")[, -c(1, 2)])
```


```{r, eval = FALSE}
set.seed(3)
num_sim <- 100

chris_vec <- rep(NA, length = num_sim)
mouth_vec <- rep(NA, length = num_sim)
pihat_vec <- rep(NA, length = num_sim)
for (index in 1:num_sim) {
  cat("Iteration:", index, "\n")
  
  num_sv <- 0
  
  while(num_sv == 0) {
    ## Simulate some data
    simdat <- poisthin(mat = t(hdat), nsamp = 10, ngene = 1000, prop_null = 0.9)
    num_sv <- num.sv(dat = t(simdat$Y), mod = simdat$X)
  }

  ## Fit MOUTHWASH
  mout <- mouthwash(Y = simdat$Y, X = simdat$X, cov_of_interest = 2, k = num_sv, detailed_output = TRUE)

  ## Calculate Chris' variance inflation factor
  xnorm <- mout$extra$rotate_out$R22^2
  znorm <- sum(mout$z2 ^ 2)
  inflate_chris <- 1 + xnorm * znorm / (nrow(simdat$X) - ncol(simdat$X))
  chris_vec[index] <- inflate_chris
  mouth_vec[index] <- mout$xi
  pihat_vec[index] <- mout$pi0
}
saveRDS(list(chris = chris_vec, mouth = mouth_vec, pi0hat = pihat_vec), "../output/chris/mvc.RDS")
```

There seems to be almost no relationship between the two variance inflation factors:
```{r}
library(tidyverse)
mvc <- readRDS("../output/chris/mvc.RDS")
mouth <- mvc$mouth
chris <- mvc$chris
qplot(mouth, chris) + theme_bw()
cor(mouth, chris)
cor(mouth, chris, method = "kendall")

boxplot(mouth, chris, names = c("mouth", "chris"))
```

Though, the MOUTHWASH inflation factors have about the same median as the Chris inflation factors, but tend to be more dispersed.
```{r}
median(mouth)
median(chris)
sd(mouth)
sd(chris)
```

It seems that MOUTHWASH isn't inflating the variances enough to get to a pi0 estimate of 0.9.
```{r}
boxplot(mvc$pi0hat)
abline(h = 0.9, lty = 2)
```

Interestingly, there doesn't seem to be a strong association between how much MOUTHWASH inflates the variances and the resulting estimates of pi0.
```{r, message=FALSE}
qplot(mouth, mvc$pi0hat) + theme_bw() +
  geom_smooth()
```


```{r}
sessionInfo()
```



