---
title: "See if confounders are picking up bimodality"
author: "David Gerard"
date: "September 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I explore if the confounders are picking up the bimodality for small $\pi_0$.

# Analysis

```{r, message=FALSE, warning=FALSE}
library(vicar)
library(sva)
library(ashr)
library(limma)
library(pROC)
load("../data/counts-allgenes-bimodal.rda")
```

```{r}
result_mat <- read.csv("../output/sc_joyce_bimodal_results.csv")
index <- 4 ## has a small pi0
result_mat$true_pi0[index]
```

```{r, cache = TRUE}
count_mat <- counts_allgenes_bimodal[[index]]$counts
Y <- log2(count_mat + 1)
X <- model.matrix(~as.factor(counts_allgenes_bimodal[[index]]$condition))
colnames(X) <- c("Intercept", "Treatment")
beta <- counts_allgenes_bimodal[[index]]$beta
is_null <- counts_allgenes_bimodal[[index]]$null

## deal with some all zero columns of counts ----------------
if (any(apply(Y, 1, sd) < 10 ^ -12)) {
  which_good <- apply(Y, 1, sd) > 10 ^ -12
  Y <- Y[which_good, ] 
  beta <- beta[which_good]
  is_null <- is_null[which_good]
  count_mat <- count_mat[which_good, ]
}

## MOUTHWASH -------------------------------------
suppressMessages(mout <- mouthwash(Y = t(Y), X = X, cov_of_interest = 2, detailed_output = TRUE))

## SVA + VOOM + LIMMA + EBAYES -------------------
trash <- capture.output(sout <- sva(dat = Y, mod = X, X[, 1, drop = FALSE]))
Xsv <- cbind(X, sout$sv)
vout <- voom(counts = count_mat, design = Xsv)
lmout <- lmFit(vout)
ebout <- ebayes(fit = lmout)
betahat <- lmout$coefficients[, 2]
sebetahat <- sqrt(ebout$s2.post) * lmout$stdev.unscaled[, 2]
degrees_freedom <- ebout$df.total[1] ## same for all indices, so choose first
suppressMessages(ashout <- ashr::ash(betahat = betahat, sebetahat = sebetahat, df = degrees_freedom))
```

## Explore MOUTHWASH confounders

The true betas are basically just -2, 0, and 2
```{r}
hist(beta)
```

The effect of the confounders is definely associated with beta.
```{r}
beta_round <- round(beta)
boxplot(mout$extra$az~beta_round)
```


```{r, eval = FALSE, echo = FALSE}
##MOUTHWASH's betahat is actually also associated with beta
mpm <- get_pm(mout)
boxplot(mpm ~ beta_round)
```


```{r, eval = FALSE, echo = FALSE}
## But ASH is a lot better
apm <- get_pm(ashout)
boxplot(apm ~ beta_round)
```


If Z2 is picking up as much of the signal as possible (because it's trying to account for deviations from unimodality), then one would expect its least squares to the true beta would be about what z2hat is. I.e.

$$
\hat{z}_2 \approx (\tilde{\alpha}^T\tilde{\alpha})^{-1}\tilde{\alpha}^T\beta
$$
Let's check:

```{r}
stupid_z2hat <- solve(crossprod(mout$extra$alpha_tilde)) %*% t(mout$extra$alpha_tilde) %*% beta
plot(mout$z2, stupid_z2hat, xlab = "Mouthwash's z2hat", ylab = "Z2 that absorbs as much beta has possible")
abline(0, 1, lty = 2)
cor(mout$z2, stupid_z2hat)
cor(mout$z2, stupid_z2hat, method = "kendall")
```

Whoa.


# Repeat with a dataset with few non-null genes

```{r}
index <- 80
result_mat$true_pi0[index]
```

```{r, cache = TRUE}
count_mat <- counts_allgenes_bimodal[[index]]$counts
Y <- log2(count_mat + 1)
X <- model.matrix(~as.factor(counts_allgenes_bimodal[[index]]$condition))
colnames(X) <- c("Intercept", "Treatment")
beta <- counts_allgenes_bimodal[[index]]$beta
is_null <- counts_allgenes_bimodal[[index]]$null

## deal with some all zero columns of counts ----------------
if (any(apply(Y, 1, sd) < 10 ^ -12)) {
  which_good <- apply(Y, 1, sd) > 10 ^ -12
  Y <- Y[which_good, ] 
  beta <- beta[which_good]
  is_null <- is_null[which_good]
  count_mat <- count_mat[which_good, ]
}

## MOUTHWASH -------------------------------------
suppressMessages(mout <- mouthwash(Y = t(Y), X = X, cov_of_interest = 2, detailed_output = TRUE))

## SVA + VOOM + LIMMA + EBAYES -------------------
trash <- capture.output(sout <- sva(dat = Y, mod = X, X[, 1, drop = FALSE]))
Xsv <- cbind(X, sout$sv)
vout <- voom(counts = count_mat, design = Xsv)
lmout <- lmFit(vout)
ebout <- ebayes(fit = lmout)
betahat <- lmout$coefficients[, 2]
sebetahat <- sqrt(ebout$s2.post) * lmout$stdev.unscaled[, 2]
degrees_freedom <- ebout$df.total[1] ## same for all indices, so choose first
suppressMessages(ashout <- ashr::ash(betahat = betahat, sebetahat = sebetahat, df = degrees_freedom))
```

```{r}
auc(predictor = get_lfdr(mout), response = is_null)
auc(predictor = get_lfdr(ashout), response = is_null)
```

```{r}
stupid_z2hat <- solve(crossprod(mout$extra$alpha_tilde)) %*% t(mout$extra$alpha_tilde) %*% beta
plot(mout$z2, stupid_z2hat, xlab = "Mouthwash's z2hat", ylab = "Z2 that absorbs as much beta has possible")
abline(0, 1, lty = 2)
cor(stupid_z2hat, mout$z2)
cor(stupid_z2hat, mout$z2, method = "kendall")
```

A little bit, but not as much.


```{r}
sessionInfo()
```
