---
title: "Try Bimodal on Bulk Data where I choose low expressed counts"
author: "David Gerard"
date: "September 21, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I try the same as code as in single cell bimodal signal [here](m_vs_svle_bimodal.html) but with bulk GTEx data.

MOUTHWASH works fine here. So I it might also have something to do with the number of zeros.

It turns out it is the zeros.

# Analysis

```{r, message=FALSE, warning=FALSE}
library(vicar)
library(sva)
library(ashr)
library(limma)
library(pROC)
library(seqgendiff)
library(tidyverse)
```

```{r}
hdat <- as.matrix(read.csv("../data/heart.csv")[, -c(1, 2)])
hrm <- rowMeans(hdat)
hdat <- hdat[hrm > 2, ]
hrm <- hrm[hrm > 2]
order_vec <- order(hrm, decreasing = FALSE)
hdat <- hdat[order_vec[1:1000], ]

load("../data/counts-allgenes-bimodal.rda")
mean(counts_allgenes_bimodal[[1]]$counts == 0)
mean(hdat == 0)
```

Signal generation function. I fiddled with the signal to get AUC comparable with that seen in the SC data.
```{r}
## signal function
gen_beta <- function(n, pi0) {
  beta <- sample(x = c(-2, 2, 0), prob = c((1 - pi0) / 2, (1 - pi0) / 2, pi0 ), replace = TRUE, size = n)
  return(beta)
}
```

```{r, cache=TRUE, eval = FALSE}
num_dat <- 100
result_mat <- matrix(NA, nrow = num_dat, ncol = 7)
colnames(result_mat) <- c("true_pi0", "ash_pi0", "mouth_pi0", "ash_auc", 
                          "mouth_auc", "ash_mse", "mouth_mse")
for (index in 1:num_dat) {
  cat("Index:", index, "\n")
  true_pi0 <- runif(1)
  
  simdat <- poisthin(mat = t(hdat), nsamp = 100, ngene = 1000, prop_null = 0, 
                     signal_fun = gen_beta, signal_params = list(pi0 = true_pi0))
  
  count_mat <- t(simdat$Y)
  Y <- log2(count_mat + 1)
  X <- simdat$X
  colnames(X) <- c("Intercept", "Treatment")
  beta <- simdat$beta
  is_null <- abs(simdat$beta) < 10 ^ -12
  
  ## deal with some all zero columns of counts ----------------
  if (any(apply(Y, 1, sd) < 10 ^ -12)) {
    which_good <- apply(Y, 1, sd) > 10 ^ -12
    Y <- Y[which_good, ] 
    beta <- beta[which_good]
    is_null <- is_null[which_good]
    count_mat <- count_mat[which_good, ]
  }
  
  ## MOUTHWASH -------------------------------------
  suppressMessages(mout <- mouthwash(Y = t(Y), X = X, cov_of_interest = 2))
  
  ## SVA + VOOM + LIMMA + EBAYES -------------------
  trash <- capture.output(sout <- sva(dat = Y, mod = X, X[, 1, drop = FALSE]))
  Xsv <- cbind(X, sout$sv)
  vout <- voom(counts = count_mat, design = Xsv)
  lmout <- lmFit(vout)
  ebout <- ebayes(fit = lmout)
  betahat <- lmout$coefficients[, 2]
  sebetahat <- sqrt(ebout$s2.post) * lmout$stdev.unscaled[, 2]
  degrees_freedom <- ebout$df.total[1] ## same for all indices, so choose first
  suppressMessages(ashout <- ashr::ash(betahat = betahat, sebetahat = sebetahat, df = degrees_freedom))
  
  ## LFDR's -----------------------------------------
  alfdr <- get_lfdr(ashout)  
  mlfdr <- get_lfdr(mout)
  
  ## posterior means --------------------------------
  apm <- get_pm(ashout)
  mpm <- get_pm(mout)

  ## Put into result_mat ---------------------------
  result_mat[index, 1] <- mean(is_null)
  result_mat[index, 2] <- get_pi0(ashout)
  result_mat[index, 3] <- get_pi0(mout)
  if (length(unique(is_null)) == 2) { ## deal with situations of all genes are null
    result_mat[index, 4] <- auc(predictor = alfdr, response = is_null)
    result_mat[index, 5] <- auc(predictor = mlfdr, response = is_null)    
  }
  result_mat[index, 6] <- mean((apm - beta) ^ 2)
  result_mat[index, 7] <- mean((mpm - beta) ^ 2)
}
write.csv(result_mat, "../output/thin_bulk_bimodal_results.csv", row.names = FALSE)
```

```{r}
result_mat <- read.csv("../output/thin_bulk_bimodal_results.csv")
```

## Plot Estimate of pi0
```{r}
as_data_frame(result_mat) %>% 
  select(true_pi0, ash_pi0, mouth_pi0) %>%
  gather(key = method, value = pi0hat, ash_pi0, mouth_pi0) ->
  longdat

pl <- ggplot(data = longdat, mapping = aes(x = true_pi0, y = pi0hat)) +
  facet_grid(.~method) +
  geom_point() +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  geom_abline(slope = 1, intercept = 0, lty = 2, col = 2, lwd = 1) +
  ylim(0, 1) +
  xlim(0, 1) +
  xlab("True Pi0")
print(pl)
```

Plot AUC boxplots

```{r}
as_data_frame(result_mat) %>%
  select(contains("auc"), true_pi0) %>%
  gather(key = "method", value = "auc", contains("auc")) ->
  longdat


ggplot(data = longdat, mapping = aes(x = method, y = auc)) +
  geom_boxplot()
```

AUC vs Pi0
```{r}
ggplot(data = longdat, mapping = aes(x = true_pi0, y = auc)) +
  facet_grid(.~method) +
  geom_point() +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  xlab("pi0") +
  ylab("auc")
```

```{r, eval = FALSE, echo = FALSE}
as_data_frame(result_mat) %>%
  select(contains("mse")) %>%
  gather(key = "method", value = "mse") ->
  longdat

ggplot(data = longdat, mapping = aes(x = method, y = mse)) +
  geom_boxplot()
```

```{r}
sessionInfo()
```
